Random Matrix Theory Tools for the Analysis of Functional Magnetic Resonance Imaging Examinations

### 2022-10-31 paper/paper/SPIE/article.tex
WIP on paper

Whether in the presence of experimental stimuli, or the relative absence, as in a resting state,
complex functional connectivity networks are ubiquitous
\citenum{bucknerBrainDefaultNetwork2008,foxCoverHumanBrain2005,gonzalez-castilloTaskbasedDynamicFunctional2018,hermundstadStructuralFoundationsRestingstate2013}.
This complexity suggests fMRI is a candidate to be studied using \textbf{Random Matrix Theory} (RMT), a set
of mathematical tools originally developed some 50 years ago to solve complex problems in nuclear
physics \citenum{guhrRandommatrixTheoriesQuantum1998a,mehtaRandomMatrices2004}. In various
physical systems where nuclei can have different possible states, precisely modeling the entirety of
the interactions between nuclei is often intractable. For such large, complex systems, the
eigenvalues (or spectra) efficiently summarize the totality of the interactions between components
of the system, and RMT describes the expected behaviour of such eigenvalues. By analysing the
spectrum statistically, complex systems can be investigated by comparison to the universal
properties predicted by RMT.

In recent years, RMT analyses have been applied to study brain functioning.  The earliest  study
demonstrated that spectra of the correlations between electroencephalographic (EEG) signals closely
resemble those of the Gaussian Orthogonal Ensemble (GOE) \citenum{sebaRandomMatrixAnalysis2003}.
Somewhat more recently, RMT was used to evaluate the quality of whole brain features extracted from
fMRI data \citenum{voultsidouFeatureEvaluationFMRI2007,verganiRestingStateFMRI2019}. RMT has also
been used in diffusion MRI to aid in the selection of the number of components to employ in
principal-component reduction analysis and denoising
\citenum{veraartDenoisingDiffusionMRI2016,verganiRestingStateFMRI2019,ulfarssonDimensionEstimationNoisy2008}.

Finally, RMT has been used in ROI-based fMRI functional-connectivity studies to investigate
differences between rest and task states \citenum{wangSpectralPropertiesTemporal2015}, between
subjects with and without ADHD \citenum{wangRandomMatrixTheory2016}, and between pain and non-pain
states \citenum{matharooSpontaneousBackpainAlters2020}. Across these three studies, the spectra of
resting or low-attention states exhibited properties closest to the GOE. These findings suggest that
certain aspects of psychological processes might be characterized, in part, by features computed
from the eigenvalues of fMRI correlation matrices, and that these features might vary in an
interpretable manner across psychological processes. If this is the case, RMT could have unique
potential in characterizing the functioning of the human brain.

@@

As our primary intention is to examine the potential utility of RMT-derived features across a wide
variety of data, and not to decisively put forward or refute any specific hypotheses relating RMT to
psychological processes, we examine a wide selection of fMRI datasets available on the
\href{https://openneuro.org/}{OpenNeuro} platform (36).  This allows for a broad assessment of the
potential of voxelwise RMT features.

However, since previous studies employing RMT for feature extraction have partially interpreted
their findings with respect to attentional processes
\citenum{wangRandomMatrixTheory2016,matharooSpontaneousBackpainAlters2020}, we have also elected to use
datasets that involve attention in some way. This includes measures of attention or vigilance, or
comparisons between a condition involving more attention and focus (e.g. a task state) or conditions
which should involve less attention or focus (e.g. resting-state, pain medication vs. none).

\subsection{Dataset: PSYCH}
The PSYCH Dataset  is detailed in \citenum{gorgolewskiHighResolution7Tesla2015}. Subjects completed a
battery of psychological measurements, and various resting-state fMRI (rs-fMRI) whole-brain 7-Tesla
scans were obtained. For the purpose of keeping with the ``attention'' theme of the analyses in this
paper, three pairings of subgroups were constructed via z-score splits of the various sum-scores of
scale items involving concepts closely related to attention.

\emph{Trait Attention}.
The PANAS-X \citenum{watsonPANASXManualPositive1994} is a 9-point Likert-type self-report measure that
queries the extent to which subjects agree particular affective terms have applied to them over a
recent time period (e.g. a few weeks). An ad hoc (i.e. face-valid) scale was constructed from the
scale items: ``attentive'', ``alert'', ``interested'', ``concentrating'', and reverse-scored items
``sluggish'', ``tired'', ``sleepy'', ``drowsy''. Subjects were then placed in the high / low ``trait
attention'' category based on a median split of the average z-score across items.

\emph{Task Attention}.
The Conjunctive Continuous Performance Task (CCPT;
\citenum{shalevConjunctiveContinuousPerformance2011} is a behavioural measure of sustained attention
where subjects must react only if they see a red square among other different coloured figures. For
this paper, a sum score was constructed by subtracting the number of false positive and false
negative responses from the number of success responses, and subjects with a z-score greater / less
than zero on this scale were placed in high / low ``task attention'' category.

\emph{Vigilance}.
The Mini New York cognition questionnaire assesses various elements of the subject's cognitive
state, e.g. ``my thoughts involved other people'', ``I thought about something positive'' (translated
from German; \citenum{gorgolewskiCorrespondenceIndividualDifferences2014}. One question assesses
alertness (``I was fully awake'' / ``war ich vollkommen wach'') and was used to create high / low
``vigilance'' conditions in the same manner as the previous two splits.

\subsubsection{Dataset 2}
Dataset 2 (referred to in figures and tables with the label ``OSTEO'') is detailed in
\citenum{tetreaultBrainConnectivityPredicts2016}. Subjects included healthy controls (``nopain''
condition) and individuals with knee osteoarthritis (``pain'' condition) treated for 2 weeks with
either placebo (``pain'' condition) or duloxetine (``duloxetine'' condition). An additional grouping was
also created with both the placebo and duloxetine data (``allpain'' condition), allowing for six
subgroup comparisons. All subjects were imaged with whole-brain rs-fMRI. Scan parameters are
documented in Table \ref{table:1}. The use of rs-fMRI and RMT for examining pain makes this dataset
complementary to \citenum{matharooSpontaneousBackpainAlters2020}.

\subsubsection{Dataset 3}
Dataset 3  (referred to here with the label ``REFLECT'', due to the ``reflective” nature of the
cognitive tasks) is described in full detail by \citenum{dupreMultiechoFMRIReplication2016}. Both
multi-echo rs- and task-fMRI scans were conducted with scan parameters summarized in Table \ref{table:2}. Tasks
involved viewing various affective images (see \citenum{sprengPatternsBrainActivity2010} for details)
and performing tasks related to autobiographical memory, prospection, or theory-of-mind reasoning.
Continuing with the theme of broadly comparing high versus low attention throughout this paper, the
different tasks were simply treated generically as ``attending'', in contrast to the non-attentive
resting-state scans. This resulted in four task state scans and one resting-state scan per subject.

In order to combine the multi-echo data into a single, usable matrix, two methods were employed. In
the ``summed'' condition, echoes were combined via summation, a method which has value in providing an
image with overall increased contrast \citenum{kunduMultiechoFMRIReview2017}. In the ``interleave''
condition, echoes were interleaved at each voxel: that is, if we denote the voxel's \(i\)th signal value
from echo \(j\) as \(t_{j,i}\), then the interleaved voxel of the new image has the time-series \([t_{1,1},
t_{2,1}, t_{3,1}, \dots, t_{1,i}, t_{2,i}, t_{3,i}, \dots, t_{1,n}, t_{2,n}, t_{3,n}]\).

@@

\subsubsection{Dataset 4}
Dataset 4 (referred to in figures and tables with the label ``PARK'' or ``PARKINSONS'') is documented and
described in detail in \citenum{madhyasthaDynamicConnectivityRest2015}. Scan parameters are documented
in Table \ref{table:1}. Subjects with non-demented Parkinson's Disease (PD) and healthy controls performed a
number of repetitions of the Attention Network Task
(ANT; \citenum{fanActivationAttentionalNetworks2005} ). Subgroup details are available in Table \ref{table:1}. As the ANT
task demands high attention, it is not as straightforward to create ``high'' or ``low'' attention
groups. Instead, since attention presumably is high and varies little in this dataset, and subgroups
instead differ by virtue of the presence or absence of PD, any results here serve as a useful
comparison point for the results from studies that should clearly involve attentional differences.
This dataset also included extensively preprocessed scans, which were included in analyses to
examine the effects of preprocessing on the extracted RMT features.

\subsubsection{Dataset 5}
Dataset 5 (referred to in figures and tables with the label ``LEARN'') is documented and described in
detail in \citenum{schapiroHumanHippocampalReplay2018}. In this study, both rs-fMRI and task fMRI scans
were available for all subjects, allowing for within-subject comparisons. The complex task involved
multiple phases, wherein subjects were exposed to named, black-and-white outlined satellite images
related by family-resemblance of the components see figure 1 of \citenum{schapiroHumanHippocampalReplay2018}. After this learning task, subjects were tested on their
memory of the various learned features. Scans corresponding to both the initial learning and
subsequent testing phase were included in the high-attention ``task'' subgroup for the purposes of
the analysis of this paper. Resting state scans were included in the low-attention ``rest''
subgroup.

@@

\begin{table}[h!]
\caption{
    ID = Identifier for paper. Subgroup = name of subgroup. Subjects = Number of subjects.
    Task = fMRI task. ANT = Attention Network Task \citenum{fanActivationAttentionalNetworks2005}
}
\label{table:2}
\small
\centering
\begin{tabular}{ l l c c c }
\hline
\textbf{ID}  & \textbf{Subgroup}  & \textbf{Subjects}  & \textbf{Task} & \textbf{Scans per Subject} \\
\hline
AGING   &  older             & 28      & resting-state         & 1 \\
        &  younger           & 34      & resting-state         & 1 \\
\hline
BILING  &  bilingual          & 59      & resting-state        & 1 \\
        &  monolingual        & 33      & resting-state        & 1 \\
\hline
DEPRESS &  depression         & 51      & resting-state        & 1 \\
        &  control            & 21      & resting-state        & 1 \\
\hline
LEARN   &  task              & 24      & learn image sketches  & 16 \\
        &  rest              & 24      & resting-state         & 2 \\
\hline
OSTEO   &  duloxetine        & 19      & taking duloxetine     & 1 \\
        &  pain              & 37      & placebo               & 1 \\
        &  nopain            & 20      & control               & 1 \\
\hline
PARK    &  Parkinson's       & 25      & ANT                   & 12 \\
        &  control           & 21      & ANT                   & 12 \\
\hline
PSYCH   &  vigilant           & 11     & resting-state        & 3 \\
        &  nonigilant         & 11     & resting-state        & 3 \\
        &  trait-attentive     & 11    & resting-state        & 3 \\
        &  trait-non-attentive & 11    & resting-state        & 3 \\
        &  task-attentive     & 11     & resting-state        & 3 \\
        &  task-nonattentive  & 11     & resting-state        & 3 \\
\hline
\end{tabular}
\end{table}

\section{Methods}

>>>>

Whether in the presence of experimental stimuli, or the relative absence, as in resting state fMRI (rs-fMRI),
complex patterns in functional connectivity networks are ubiquitous
\cite{bucknerBrainDefaultNetwork2008,foxCoverHumanBrain2005,gonzalez-castilloTaskbasedDynamicFunctional2018,hermundstadStructuralFoundationsRestingstate2013}.
This complexity suggests fMRI is a candidate to be studied using \textbf{Random Matrix Theory} (RMT), a set
of mathematical tools originally developed some 50 years ago to solve complex problems in nuclear
physics \cite{guhrRandommatrixTheoriesQuantum1998a,mehtaRandomMatrices2004}

In various
physical systems where nuclei can have different possible states, precisely modeling the entirety of
the interactions between nuclei is often intractable. For such large, complex systems, the
eigenvalues (or spectra) efficiently summarize the totality of the interactions between components
of the system, and RMT describes the expected behaviour of such eigenvalues. By analysing the
spectrum statistically, complex systems can be investigated by comparison to the universal
properties predicted by RMT.

\subsection{RMT and Neurobiological Signals}

In recent years, RMT analyses have been applied to study brain functioning. The earliest study
demonstrated that spectra of the correlations between electroencephalographic (EEG) signals closely
resemble those of the Gaussian Orthogonal Ensemble (GOE) \cite{sebaRandomMatrixAnalysis2003}.
Somewhat more recently, RMT was used to evaluate the quality of whole brain features extracted from
fMRI data \cite{voultsidouFeatureEvaluationFMRI2007,verganiRestingStateFMRI2019}. RMT has also
been used in diffusion MRI to aid in the selection of the number of components to employ in
principal-component reduction analysis and denoising
\cite{veraartDenoisingDiffusionMRI2016,verganiRestingStateFMRI2019,ulfarssonDimensionEstimationNoisy2008}.

Finally, RMT has been used in ROI-based fMRI functional-connectivity studies to investigate
differences between rest and task states \cite{wangSpectralPropertiesTemporal2015}, between
subjects with and without ADHD \cite{wangRandomMatrixTheory2016}, and between pain and non-pain
states \cite{matharooSpontaneousBackpainAlters2020}. Across these three studies, the spectra of
resting or low-attention states exhibited properties closest to the GOE. These findings suggest that
certain aspects of psychological processes might be characterized, in part, by features computed
from the eigenvalues of fMRI correlation matrices, and that these features might vary in an
interpretable manner across psychological processes. If this is the case, RMT could aid
in understanding the functioning of the human brain.

\subsection{Eigenfeatures}

The basic insight of RMT is thus that the \textit{eigenvalues alone} may provide interesting
information about highly complex Neurobiological systems. However, in practice, the eigenvalues computed from
empirically observed data are noisy. RMT is statistical in nature, describing only the
\textit{expected} behaviour of the system spectra. To this end, a number of summary statistics or
``spectral observables'' \cite{mehtaRandomMatrices2004} can be computed from empirically observed
spectra, with these spectral observables sometimes being better suited for further analysis or
comparison to theory. Two such summary statistics that have been somewhat popular are the
\textit{spectral rigidity} and \textit{level number variance} (``rigidity'' and ``level variance''
for short; details in Section \ref{sect:methods}).

However, from a predictive standpoint, these summary statistics may eliminate predictively-useful
information. If the basic RMT insight is that the eigenvalues are useful features to understand a
system, then those eigenvalues alone (or other simple transformations of them) ought also to be
predictively useful. We examine a number of such \textit{eigenfeatures} in this study.

\subsection{Limitations of Previous Work}

However, these previous studies took largely descriptive / explanatory approaches, noting only
differences in RMT across subgroups and/or conditions, in many cases, with no formal statistical
model or testing of the significance and robustness of these differences. No code was provided to
allow the reproduction of the results, and the extraction of the spectral rigidity and level
variance is non-trivial, and computationally demanding. Computation of these spectral observables
requires an \textit{unfolding}
procedure\cite{guhrRandommatrixTheoriesQuantum1998a,mehtaRandomMatrices2004}, which is,
unfortunately, a poorly-documented fitting procedure that requires making a number of
difficult-to-justify decisions, and these decisions can dramatically impact RMT conclusions
\cite{abul-magdUnfoldingSpectrumChaotic2014,abueleninSpectralUnfoldingChaotic2018,fossionRandommatrixSpectraTime2013,abueleninEffectUnfoldingSpectral2012,moralesImprovedUnfoldingDetrending2011}.
On top of the flexibility in the implementation and application of RMT to empirical data, there
is also the flexibility introduced by the choices in the complex preprocessing pipelines of fMRI\cite{parkerBenefitSliceTiming2019}.

What is missing is a rigorous, systematic, reproducible investigation of the *predictice* value of
RMT metrics across a wide variety of data, RMT analytic choices, choice of eigenfeature, and choice
of preprocessing. The current study aims to rectify this.

@@

To examine the potential utility of RMT-derived features across a wide
variety of data, and not to decisively put forward or refute any specific hypotheses relating RMT to
psychological processes, we examine a wide selection of fMRI datasets available on the
\href{https://openneuro.org/}{OpenNeuro} platform (36).  This allows for a broad assessment of the
potential of voxelwise RMT features.

However, since previous studies employing RMT for feature extraction have partially interpreted
their findings with respect to attentional processes
\cite{wangRandomMatrixTheory2016,matharooSpontaneousBackpainAlters2020}, we have also elected to use
datasets that involve attention in some way. This includes measures of attention or vigilance, or
comparisons between a condition involving more attention and focus (e.g. a task state) or conditions
which should involve less attention or focus (e.g. resting-state, pain medication vs. none).

\subsection{Dataset: PSYCH}
The PSYCH Dataset  is detailed in \cite{gorgolewskiHighResolution7Tesla2015}. Subjects completed a
battery of psychological measurements, and various resting-state fMRI (rs-fMRI) whole-brain 7-Tesla
scans were obtained. For the purpose of keeping with the ``attention'' theme of the analyses in this
paper, three pairings of subgroups were constructed via z-score splits of the various sum-scores of
scale items involving concepts closely related to attention.

\emph{Trait Attention}.
The PANAS-X \cite{watsonPANASXManualPositive1994} is a 9-point Likert-type self-report measure that
queries the extent to which subjects agree particular affective terms have applied to them over a
recent time period (e.g. a few weeks). An ad hoc (i.e. face-valid) scale was constructed from the
scale items: ``attentive'', ``alert'', ``interested'', ``concentrating'', and reverse-scored items
``sluggish'', ``tired'', ``sleepy'', ``drowsy''. Subjects were then placed in the high / low ``trait
attention'' category based on a median split of the average z-score across items.

\emph{Task Attention}.
The Conjunctive Continuous Performance Task \cite{shalevConjunctiveContinuousPerformance2011} (CCPT)
 is a behavioural measure of sustained attention
where subjects must react only if they see a red square among other different coloured figures. For
this paper, a sum score was constructed by subtracting the number of false positive and false
negative responses from the number of success responses, and subjects with a z-score greater / less
than zero on this scale were placed in high / low ``task attention'' category.

\emph{Vigilance}.
The Mini New York cognition questionnaire assesses various elements of the subject's cognitive
state, e.g. ``my thoughts involved other people'', ``I thought about something positive'' (translated
from German; \cite{gorgolewskiCorrespondenceIndividualDifferences2014}. One question assesses
alertness (``I was fully awake'' / ``war ich vollkommen wach'') and was used to create high / low
``vigilance'' conditions in the same manner as the previous two splits.

\subsubsection{Dataset 2}
Dataset 2 (referred to in figures and tables with the label ``OSTEO'') is detailed in
\cite{tetreaultBrainConnectivityPredicts2016}. Subjects included healthy controls (``nopain''
condition) and individuals with knee osteoarthritis (``pain'' condition) treated for 2 weeks with
either placebo (``pain'' condition) or duloxetine (``duloxetine'' condition). An additional grouping was
also created with both the placebo and duloxetine data (``allpain'' condition), allowing for six
subgroup comparisons. All subjects were imaged with whole-brain rs-fMRI. Scan parameters are
documented in Table \ref{table:1}. The use of rs-fMRI and RMT for examining pain makes this dataset
complementary to \cite{matharooSpontaneousBackpainAlters2020}.

\subsubsection{Dataset 3}
Dataset 3  (referred to here with the label ``REFLECT'', due to the ``reflective” nature of the
cognitive tasks) is described in full detail by \cite{dupreMultiechoFMRIReplication2016}. Both
multi-echo rs- and task-fMRI scans were conducted with scan parameters summarized in Table \ref{table:2}. Tasks
involved viewing various affective images (see \cite{sprengPatternsBrainActivity2010} for details)
and performing tasks related to autobiographical memory, prospection, or theory-of-mind reasoning.
Continuing with the theme of broadly comparing high versus low attention throughout this paper, the
different tasks were simply treated generically as ``attending'', in contrast to the non-attentive
resting-state scans. This resulted in four task state scans and one resting-state scan per subject.

In order to combine the multi-echo data into a single, usable matrix, two methods were employed. In
the ``summed'' condition, echoes were combined via summation, a method which has value in providing an
image with overall increased contrast \cite{kunduMultiechoFMRIReview2017}. In the ``interleave''
condition, echoes were interleaved at each voxel: that is, if we denote the voxel's \(i\)th signal value
from echo \(j\) as \(t_{j,i}\), then the interleaved voxel of the new image has the time-series \([t_{1,1},
t_{2,1}, t_{3,1}, \dots, t_{1,i}, t_{2,i}, t_{3,i}, \dots, t_{1,n}, t_{2,n}, t_{3,n}]\).

@@

\subsubsection{Dataset 4}
Dataset 4 (referred to in figures and tables with the label ``PARK'' or ``PARKINSONS'') is documented and
described in detail in \cite{madhyasthaDynamicConnectivityRest2015}. Scan parameters are documented
in Table \ref{table:1}. Subjects with non-demented Parkinson's Disease (PD) and healthy controls performed a
number of repetitions of the Attention Network Task
(ANT; \cite{fanActivationAttentionalNetworks2005} ). Subgroup details are available in Table \ref{table:1}. As the ANT
task demands high attention, it is not as straightforward to create ``high'' or ``low'' attention
groups. Instead, since attention presumably is high and varies little in this dataset, and subgroups
instead differ by virtue of the presence or absence of PD, any results here serve as a useful
comparison point for the results from studies that should clearly involve attentional differences.
This dataset also included extensively preprocessed scans, which were included in analyses to
examine the effects of preprocessing on the extracted RMT features.

\subsubsection{Dataset 5}
Dataset 5 (referred to in figures and tables with the label ``LEARN'') is documented and described in
detail in \cite{schapiroHumanHippocampalReplay2018}. In this study, both rs-fMRI and task fMRI scans
were available for all subjects, allowing for within-subject comparisons. The complex task involved
multiple phases, wherein subjects were exposed to named, black-and-white outlined satellite images
related by family-resemblance of the components see figure 1 of \cite{schapiroHumanHippocampalReplay2018}. After this learning task, subjects were tested on their
memory of the various learned features. Scans corresponding to both the initial learning and
subsequent testing phase were included in the high-attention ``task'' subgroup for the purposes of
the analysis of this paper. Resting state scans were included in the low-attention ``rest''
subgroup.

@@

\begin{table}[h!]
\caption{
    ID = Identifier for paper. Subgroup = name of subgroup. Subjects = Number of subjects.
    Task = fMRI task. ANT = Attention Network Task \cite{fanActivationAttentionalNetworks2005}
}
\label{table:2}
\small
\centering
\begin{tabular}{ l l c c c }
\hline
\textbf{ID}  & \textbf{Subgroup}  & \textbf{Subjects}  & \textbf{Task} & \textbf{Scans per Subject} \\
\hline
AGING   &  older             & 28      & resting-state         & 1 \\
        &  younger           & 34      & resting-state         & 1 \\
\hline
BILING  &  bilingual          & 59      & resting-state        & 1 \\
        &  monolingual        & 33      & resting-state        & 1 \\
\hline
DEPRESS &  depression         & 51      & resting-state        & 1 \\
        &  control            & 21      & resting-state        & 1 \\
\hline
LEARN   &  task              & 24      & learn image sketches  & 16 \\
        &  rest              & 24      & resting-state         & 2 \\
\hline
OSTEO   &  duloxetine        & 19      & taking duloxetine     & 1 \\
        &  pain              & 37      & placebo               & 1 \\
        &  nopain            & 20      & control               & 1 \\
\hline
PARK    &  Parkinson's       & 25      & ANT                   & 12 \\
        &  control           & 21      & ANT                   & 12 \\
\hline
PSYCH   &  vigilant           & 11     & resting-state        & 3 \\
        &  nonigilant         & 11     & resting-state        & 3 \\
        &  trait-attentive     & 11    & resting-state        & 3 \\
        &  trait-non-attentive & 11    & resting-state        & 3 \\
        &  task-attentive     & 11     & resting-state        & 3 \\
        &  task-nonattentive  & 11     & resting-state        & 3 \\
\hline
\end{tabular}
\end{table}

\section{Methods}
\label{sect:methods}

Given empirically observed eigenvalues \(\Lambda\), the spectral rigidity, \(\Delta_3(L)\), is
calculated for any positive real value \(L < \max(\Lambda)\) as:
\[
\Delta_3(L) = \left \langle \min_{A,B} \frac{1}{L} \int_c^{c+L} \left(  \eta(\lambda) -A \lambda - B \right)^2 \right \rangle_c
\]
where \(\eta(\lambda)\) is the number of unfolded eigenvalues less than or equal to \(\lambda\),
\(\langle \cdot \rangle_c\) denotes the average with respect to all starting points \(c\), and where
\(A\) and \(B\) denote the slope and intercept, respectively, of the least squares fit of a straight
line to \(\eta(\lambda)\) on \([c, c+L]\) \cite{guhrRandommatrixTheoriesQuantum1998a}. The spectral
rigidty for a value \(L\) is thus the \textit{average nonlinearity}

The level
number variance, \(\Sigma^2(L)\), is closely related to the spectral rigidity
\cite{mehtaRandomMatrices2004}, and is calculated as:

\[
\Sigma^2(L) = \left\langle \eta^2(L, c) \right\rangle_c - \left\langle \eta(L, c) \right\rangle^2_c
\]

where \(\eta(L, c)\) is the number of unfolded eigenvalues in \([c, c+ L]\), and where \(c\), \(L\),
and \(\langle \cdot \rangle_c\) are as above \cite{guhrRandommatrixTheoriesQuantum1998a}. Both the
spectral rigidity and level number variance are also computed and investigated in this study.
